---
title: "ADS503 Final Project"
output:
  word_document: default
  pdf_document: default
---



  Attribute Information:     
     Attribute:                Attribute Range:
     ------------------        -----------------------------------------------
  1. symboling(risk-Rating):              -3, -2, -1, 0, 1, 2, 3.
  2. normalized-losses:        continuous from 65 to 256.
  3. make:                     alfa-romero, audi, bmw, chevrolet, dodge, honda,
                               isuzu, jaguar, mazda, mercedes-benz, mercury,
                               mitsubishi, nissan, peugot, plymouth, porsche,
                               renault, saab, subaru, toyota, volkswagen, volvo
  4. fuel-type:                diesel, gas.
  5. aspiration:               std, turbo.
  6. num-of-doors:             four, two.
  7. body-style:               hardtop, wagon, sedan, hatchback, convertible.
  8. drive-wheels:             4wd, fwd, rwd.
  9. engine-location:          front, rear.
 10. wheel-base:               continuous from 86.6 120.9.
 11. length:                   continuous from 141.1 to 208.1.
 12. width:                    continuous from 60.3 to 72.3.
 13. height:                   continuous from 47.8 to 59.8.
 14. curb-weight:              continuous from 1488 to 4066.
 15. engine-type:              dohc, dohcv, l, ohc, ohcf, ohcv, rotor.
 16. num-of-cylinders:         eight, five, four, six, three, twelve, two.
 17. engine-size:              continuous from 61 to 326.
 18. fuel-system:              1bbl, 2bbl, 4bbl, idi, mfi, mpfi, spdi, spfi.
 19. bore(diameter of each cylinder):   continuous from 2.54 to 3.94.
 20. stroke:                   continuous from 2.07 to 4.17.
 21. compression-ratio:        continuous from 7 to 23.
 22. horsepower:               continuous from 48 to 288.
 23. peak-rpm:                 continuous from 4150 to 6600.
 24. city-mpg (gal):           continuous from 13 to 49.
 25. highway-mpg (gal):        continuous from 16 to 54.
 26. price:                    continuous from 5118 to 45400.


# 1 Data Restructuring

# A. What is proportion and pattern of missing values ?
# B. is missigness mechanism of Data is MCAR ?

```{r warning=FALSE, paged.print=FALSE}
# Required Libraries for project
library("mice")
library("VIM")
library("tidyverse")
library(visdat)
library(lattice)
library(ggpubr)
library("caret")
library(randomForest)
library("plyr")

df <- read.csv("C:/Users/emrek/Desktop/Ads 503/Data_Car.csv")

head(df)
str(df)

vis_miss(df) # shows missign values proportion both total and each column

summary(df)

colSums(is.na(df))

miss_plot <-aggr(df, col=mdc(1:2), numbers=TRUE, sortVars=TRUE, labels=names(df), cex.axis=.7, gap=3, xlab=c("columns"),ylab=c("Proportion of missingness","Missingness Pattern"))


# Multiple imputation
df_mice_imputes = mice(df, m=5, maxit = 40)

# Goodness of fit
densityplot(df_mice_imputes)
# Imputation
Imputed_data <- complete(df_mice_imputes,1)


#  Volume of cars 

Imputed_data$volume <- Imputed_data$X11 *Imputed_data$X12 * Imputed_data$X13
# get rid of with,length and height

#
Imputed_data <- select(Imputed_data,-X11,-X12,-X13)

Imputed_data$X1 = as.factor(Imputed_data$X1) # X1 symbol rating is discrete

# write to csv to reuse 
write.csv(Imputed_data,"C:/Users/emrek/Desktop/Ads 503/Imputed.csv",row.names = FALSE)



```




```{r}

# 2) Exploratory Analysis
# Descriptive Analyisis 

options(scipen = 999)
data <- read.csv("C:/Users/emrek/Desktop/Ads 503/Imputed.csv")


numeric_columns <- dplyr::select_if(data, is.numeric)

list_names = names(numeric_columns) # list of column names

# X2
boxplot(data$X2,xlab = "normalized-losses(X2)") # No outliers

print(paste("Median of normalized-losses:", round(median(data$X2),3),"IQR:",round(IQR(data$X2),3)))

# X10
boxplot(data$X10,xlab = "Wheel-base(X10)")
print(paste("Median of  wheel-base:", median(data$X10) ,"Measure of Variance:",round(IQR(data$X10),1)))

# X14
boxplot(data$X14,xlab = "curb-weight(X14)")
print(paste("Median of  curb-weight:", round(median(data$X14),2) ,"IQR:",round(IQR(data$X14),2)))

#x17
boxplot(data$X17,xlab = "engine-size(X17)")
print(paste("Median of  engine-size:", median(data$X17) ,"Measure of Variance:",round(IQR(data$X17),1)))

# X19
boxplot(data$X19,xlab = "bore(X19)")
print(paste("Median of  bore:", round(median(data$X19),2) ,"IQR:",round(IQR(data$X19),2)))
# X20
boxplot(data$X20,xlab = "stroke(X20)")
print(paste("Median of  stroke:", median(data$X20) ,"Measure of Variance:",round(IQR(data$X20),1)))
# X21
boxplot(data$X21,xlab = "compression-ratio(X21)")
print(paste("Median of  compression-ratio:", median(data$X21) ,"Measure of Variance:",round(IQR(data$X21),1)))
# X22
boxplot(data$X22,xlab = "horsepower(X22)")
print(paste("Median of  horsepower:", median(data$X22) ,"Measure of Variance:",round(IQR(data$X22),1)))
  # X23
boxplot(data$X23,xlab = "peak-rpm(X23)")
print(paste("Median of  peak-rpm:", median(data$X23) ,"Measure of Variance:",round(IQR(data$X23),1)))
# X24
boxplot(data$X24,xlab = "city-mpg (gal)")
print(paste("Median of  city-mpg (gal):", median(data$X24) ,"Measure of Variance:",round(IQR(data$X24),1)))
# X25
boxplot(data$X25,xlab = "highway-mpg (gal)")
print(paste("Median of  highway-mpg (gal):", median(data$X25) ,"Measure of Variance:",round(IQR(data$X25),1)))

# X26(price)
boxplot(data$X26,xlab = "Price(X26)")
print(paste("Median of  Price:", median(data$X26) ,"Measure of Variance:",round(IQR(data$X26),1)))
# volume
boxplot(data$volume,xlab = "Volume")
print(paste("Median of  Volume:", median(data$volume) ,"Measure of Variance:",round(IQR(data$volume),1)))
```


```{r}
#not to use exponential notation (e.g. e+10)

options(scipen = 999)



# What is the Distribution of price (response variable) ? 



ggplot(data,aes(x=X26))+
  geom_histogram(fill="white",col="black",aes(y=..density..)) +
  geom_density(fill = "orange", alpha = .2)+
  labs(x="price") + theme_classic()


# What is the Distribution of Volume?


ggplot(data,aes(x=volume))+
  geom_histogram(fill="white",col="black",aes(y=..density..)) +
  geom_density(fill = "orange", alpha = .2)+
  labs(x="Volume") + theme_classic()

# What is the Distribution of Engine Size ?


ggplot(data,aes(x=X17))+
  geom_histogram(fill="white",col="black",aes(y=..density..)) +
  geom_density(fill = "orange", alpha = .2)+
  labs(x="Engine Size") + theme_classic()

# What is the Distribution of Horse Power ?


ggplot(data,aes(x=X22))+
  geom_histogram(fill="white",col="black",aes(y=..density..)) +
  geom_density(fill = "orange", alpha = .2)+
  labs(x="Horsepower") + theme_classic()


# How does risk rating affect average prices of cars?
data$X1 <- as.factor(data$X1)



data  %>% group_by(X1) %>% summarise(mean_price = mean(X26)) %>% 
  ggplot(aes(mean_price,fct_reorder(X1,mean_price))) +
  geom_col(show.legend = FALSE,fill = "cadetblue4") +
  labs(y= " Risk Rating",x= "Price") +
  theme_classic()
   

# Which car brand is the most expensive ?


data  %>% group_by(X3)  %>% summarise(mean_price = mean(X26))  %>% 
  ggplot(aes(mean_price,fct_reorder(X3,mean_price),fill = X3)) + 
  geom_col(show.legend = FALSE) + 
  labs(y= "Brand of Cars",x= "Price",title = "Price vs Brands") +
  theme_classic()

  # What is frequency (distribution of cars ) of Fuel types 
count(data,'X4')

#  How does price change according to fuel type?

ggboxplot(data, x = "X4", y = "X26", 
          color = "X4",
          ylab = "Price", xlab = "Fuel Type")


# What is the frequency of number of doors?

count(data,'X6')

# How does price change according to number of cylender and number of doors ?
bwplot(reorder(X16,X26) ~ (X26)|X6, data = data,
  xlab = "Price", ylab = "Number of Cylinders",
  main = "Price According to Number of Doors and Number of Cylinders"
)
 # How does price change according to number of doors and  fuel system ?
bwplot(reorder(factor(X18),X26) ~ (X26)|(X6), data = data,
  xlab = "Price", ylab = "Fuel System",
  main = "Price According to Number of Doors and Fuel System"
)

 # How does fuel consumption in city  effect prices?
ggscatter(data,x = "X26", y = "X24",
          add ="reg.line",
          conf.int = TRUE,
          cor.coef = TRUE,
          cor.method = "spearman",
          xlab="Price",
          ylab = "city-mpg (gal) "
          )

# How does engine size effect fuel conspumption of cars using different fuel type ?

ggplot(data,aes(x=X24,X17,color = X4)) + geom_point() + 
   geom_smooth(method ="lm") +
  labs(x="city-mpg",y="Engine Size",color ="Fuel type") + theme_classic()

# Is there a relationship between price and engine size ?



ggscatter(data,x = "X17", y = "X26",
          add ="reg.line",
          conf.int = TRUE,
          cor.coef = TRUE,
          cor.method = "spearman",
          xlab="Engine Size",
          ylab = "Price",
          )
# Is there a relationship between volume and price ?
ggscatter(data,x = "volume", y = "X26",
          add ="reg.line",
          conf.int = TRUE,
          cor.coef = TRUE,
          cor.method = "spearman",
          ylab = "Price",
          xlab="Volume",
          )


# Does the car volume and door number affect the price?

ggplot(data = data,aes(y=X26,x=volume,color=X6,shape = X6)) + 
  geom_point() +  geom_smooth(method ="lm") +
  labs(y = "price",x = "Volume",
    title = "Price According to Number of Doors and  Volume",
    shape = "No of Doors",color = "No of Doors") +
  theme_classic()




```



```{r}

# 3 Inferential Analysis

options(scipen = 999)

# Is there a difference between gas and diesel car in terms of price ?

# I.Hypothesis 
# H0: There is no difference fuel types of car in terms of price 
# H1: Disel cars have higher price (disel > gas)

diesel = as.vector(data$X26[data$X4 == "diesel"])
gas = as.vector(data$X26[data$X4 == "gas"])

# Assumptions

shapiro.test(diesel)
shapiro.test(gas)


# Normality was refused 
# Non Parametric  Statistical Test

# Mann-Whitney U test

wilcox.test(diesel,gas,alternative = "greater")

# Ho was rejected since p<0.05 so we can conclude that price of diesel car greater than gasoline(gas) car

#  Is there difference between body types in terms of price ?

# # II.Hypothesis 
# HO: Body types have no impacts on price
# H1: Body types have imapcts on price

# Normality Assumption

shapiro.test(data$X26[data$X7 =="hardtop"]) # normality ok
shapiro.test(data$X26[data$X7 =="wagon"]) #  normality  was rejected
shapiro.test(data$X26[data$X7 =="sedan"]) # normality was rejected
shapiro.test(data$X26[data$X7 =="hatchback"]) # normality was rejected
shapiro.test(data$X26[data$X7 =="convertible"]) # normality ok 

# Totality normality was rejected

# Non parametric unpaired test Kruskall - Wallis

kruskal.test(X26~X7,data = data) # Ho was rejected

# Acording to Kruskall test one of body types have significant differences impacts on price

# Multiple Comparision Post Test  "Pairwise Wilcox Test

pairwise.wilcox.test(data$X26,data$X7, p.adjust.method = "BH")

#  According to pairwise comparision hardtop - convertible, hardtop,hatchback, hatch-convertible  
# hatch- sedan and hatch- wagons are significantly different.

# III.Hypothesis
# H0: There is neaning full relationship (correlation) between volume and price
# H1: There is a meaningfull relationship between volume and price

# Normality Assumption

shapiro.test(data$volume) # normality was rejected

shapiro.test(data$X26) # normality was rejected

# Non Parametric Test (Spearman Cor. Test)

cor.test(data$volume,data$X26,method = "spearman") # rho = 0.70 (correlation coefficent)
# we can conclute that that there is a meaningful correlation between two variables 
#and it is post.strong correlation

# IV Hypothesis
# H0: There is no difference fuel type of car in terms of city-mpg fuel conspumtion
# H1: There is a significant difference between two types of fuelt types in terms of fuel consumption

disel_f <- as.vector(data$X24[data$X4 == "diesel"])
gas_f <- as.vector(data$X24[data$X4 == "gas"])

# Normality Assumption

shapiro.test(disel_f) # normality was rejected
shapiro.test(gas_f)

#  Non Parametric Test - Mann-Whitney U test

wilcox.test(disel_f,gas_f) # They are signficiantly different

# Which one is greater?

ggboxplot(data, x = "X4", y = "X24", 
          color = "X4",
          ylab = "City-mile per galoon(mpg)", xlab = "Fuel Type")

# According to graph in 1985 diesel cars spend less fuel in comparision to gasoline cars

# Is Number of doors important for price ?

# V Hypothesis
# H0: There is no difference between price of two doors and four doors cars
# H1: There is a difference between price of cars according to number of doors 

two  <- as.vector(data$X26[data$X6 == "two"])
four <- as.vector(data$X26[data$X6 == "four"])


# Normality Assumption

shapiro.test(two) # normality was rejected
shapiro.test(four) # normality was rejected


# Non Parametric  Statistical Test

# Mann-Whitney U test

wilcox.test(two,four) 
# Since p is greater than 0.05 Ho cannot rejected. There is no difference between no of doors

# VI. Hypothesis


# Normality Assumption was rejected since price is not normal


# Non parametric unpaired test Kruskall - Wallis

kruskal.test(X26~X1,data = data) 

pairwise.wilcox.test(data$X26,data$X1, p.adjust.method = "BH")

```




```{r echo=FALSE}

# For model we do not need make column (x3)
options(scipen = 999)

df1 <- data

df1 <- dplyr::select(df1,-X3)

#  Transformation of categoric variables to discrete ones

must_convert<-sapply(df1,is.factor) # logical vector telling if a variable needs to be displayed as numeric
M<-sapply(df1[,must_convert],unclass) # data.frame of all categorical variables now displayed as numeric

data_model <-cbind(df1[,!must_convert],M) # complete data.frame with all variables put together


# Convert discrete variables to factor


data_model$X4 = as.factor(data_model$X4)
data_model$X5 = as.factor(data_model$X5)
data_model$X6 = as.factor(data_model$X6)
data_model$X7 = as.factor(data_model$X7)
data_model$X8 = as.factor(data_model$X8)
data_model$X9 = as.factor(data_model$X9)
data_model$X15 = as.factor(data_model$X15)
data_model$X16 = as.factor(data_model$X16)
data_model$X18 = as.factor(data_model$X18)

str(data_model)

# Splitting data as test and traing with proportion 75%


set.seed(15)
trainIndex <- createDataPartition(data_model$X26, p = .75, 
                                  list = FALSE, 
                                  times = 1)

train <- data_model[ trainIndex,] # 0.75 for train
test  <- data_model[-trainIndex,] # 0.25 for test

# Creation Random forest model without tuning and cv

# mtry: Number of variables randomly sampled as candidates at each split.
# ntree: Number of trees to grow.

#  For this method mtry is default (ncol(train)/3-1)

set.seed(15)
rf_model = randomForest(X26~., data = train,importance = TRUE)
rf_model

importance(rf_model)
varImpPlot(rf_model)

RFtrain_pred = predict(rf_model, newdata = train)

# Train Error
postResample(RFtrain_pred, train$X26)


# Test error
defaultSummary(data.frame(obs = test$X26,pred = predict(rf_model, test)))

set.seed(16)
# repeated K-FOLD cross validation
ctrl <- trainControl(method = "repeatedcv", 
                     number = 10,
                     repeats  = 3,
                     search = "grid")

tune_grid <- expand.grid(mtry = c(1:22))

rf_tune <- train(X26~.,
                 data= train,
                  method = "rf",
                  tuneGrid = tune_grid,
                  trControl = ctrl

)
# pamater tunning with cross-validation

rf_tune
plot(rf_tune,xlab ="mtry")

rf_tune$results %>% filter(mtry ==as.numeric(rf_tune$bestTune))

# test error with best parameter 
defaultSummary(data.frame(obs = test$X26, 
                            pred = predict(rf_tune, test)))




# final model
rf_final <- rf_tune$finalModel

importance( rf_final)
plot(rf_final) # best mtry effect of n_tree 

# Method of Tunning for n_tree parameter but it requires lots of time

model_listesi <- list()

for (ntree in c(50,60,70,80,90,100)) {
  
  set.seed(123)
  
  fit <- train(X26~.,
                  data = train,
                  method = "rf",
                  tuneGrid = tune_grid,
                  trControl = ctrl, 
                  ntree = ntree)

  key <- toString(ntree)
  model_listesi[[key]] <- fit
  
}
sonuclar <- resamples(model_listesi)
summary(sonuclar) 




```

